\documentclass[openany,letterpaper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}
\usepackage{tgadventor}
\usepackage{XCharter}
\usepackage{titling}
\usepackage{FiraMono}
\usepackage{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage[margin=.9in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{mdframed}
\usepackage{lastpage}
\pagenumbering{gobble}
% FOOTER

\setlist{nosep}
\author{Connor Duncan}
\title{Curriculum Vitae}
\date{\today}
\setlength\parindent{0pt}

% new title command
\renewcommand{\maketitle}{\begin{minipage}[t]{\textwidth}
    \par{\centering{\LARGE{\theauthor}}\par}
    \centering\small\color{gray}

     \href{mailto:cdunca9@uic.edu}{cdunca9@uic.edu}
    \end{minipage}
    \\
}
% changing title formatting
\titleformat{\section}{\sffamily\scshape\raggedright}{}{0em}{}[{\titlerule[1pt]}]
\titlespacing{\section}{0pt}{3pt}{7pt}
\titleformat{\subsection}{\bfseries}{}{2em}{}
\titlespacing{\subsection}{0pt}{7pt}{7pt}

% itemize
\renewcommand\leftmargini{1.5em}
\renewcommand\labelitemi{$\diamond$}
\newcommand{\Cal}{
    University of California, Berkeley}

\newcommand{\entry}[4]{ % DATE, POSITION, BUSINESS, DUTIES.
    \begin{minipage}[t]{0.15\textwidth}
        \scshape#1
    \end{minipage}
    \vline\hspace{1em}
    \begin{minipage}[t]{0.83\textwidth}
        \raggedright
        {\large\scshape#3}\\
        \vspace{0.3em}
        \vbox{\normalsize\textit{#2}#4}
    \end{minipage}
}

\newcommand{\skill}[2]{ % TYPE, LIST
    \begin{minipage}[t]{0.15\textwidth}
        \scshape#1
    \end{minipage}
    \vline\hspace{1em}
    \begin{minipage}[t]{0.83\textwidth}
        \raggedright
        #2
    \end{minipage}
}
\begin{document}
\maketitle
\vspace{1em}

Eager, motivated and analytical mathematician with experience in 
optimization, data science, software development and physics.
Seeking to apply my passion for math to techonology in an
environment that fosters professional and intellectual growth.

\section{Education}
\entry{2022--}{M.S. in Applied Mathematics}{University of Illinois at Chicago}{ || 3.86/4.0 GPA}

\entry{2017--2021}{B.A. in Physics; B.A. in Applied Mathematics}{\Cal}{ || 3.366/4.0 GPA}

\section{Experience}
\entry{Summer 2023}{ETF Portfolio Management Intern}{Invesco}{
    \begin{itemize}
        \item Used Dash, AG-Grid and Snowflake to develop quantitative tools 
            tracking rebalance schedules, index
            changes and trading fees. 
        \item Answered ad-hoc requests as required.
    \end{itemize}
}

\entry{2022--}{Teaching Assistant (Algebra \& Linear Algebra)}{University of Illinois at Chicago}{
} 

\entry{Summer 2020}{SULI Intern}
{National Renewable Energy Laboratory}
{
    \begin{itemize}
        \item Developed JavaScript application visualizing the
projected socioeconomic impact of environmental policy changes in the Los Angeles 
Area
    \end{itemize}
}

\entry{2018--2020}{Research Assistant, Pyle Group}{\Cal; SLAC}{
    \begin{itemize}
        \item Assisted with installation of Helium Dilution Refrigerator
        \item Assisted with Linux server administration (bare metal) 
            and DAQ Dashboard development.
    \end{itemize}
}



\section{Technical Skills}
\skill{Python}{
    \begin{itemize}
            \item WSGI implementation using
                Flask, including streaming applications.
    \end{itemize}
}
\skill{JavaScript}{
    \begin{itemize}
        \item React, ag-grid, visx, node.
    \end{itemize}
}
\skill{Linux}{
    \begin{itemize}
        \item Enterprise distributions such as RHEL and Debian. 
        \item systemd, administering apache/nginx, sshd, Coreutils, network configuration.
        \item Daily Arch Linux user since 2018.
    \end{itemize}
}

\skill{Additional}{
    \begin{itemize}
        \item SQL, Redis, Git (incl. Bitbucket Pipelines), scripting (zsh), 
            rsync, MATLAB, C, LabView, \LaTeX, HTML/CSS, vim, Excel, Bloomberg.
    \end{itemize}
}

\section{Other}
\skill{Spanish}{
    \begin{itemize}
        \item Conversational speaker.
    \end{itemize}
}
\skill{Interests}{
    \begin{itemize}
        \item {\scshape At Work}: 
            Numerical Methods in Partial Differential Equations, Linux Administration,
            Data Science, Performance Optimization.
        \item {\scshape After Work}: Piano, 
            Traveling, Strategy Games (e.g. Stellaris).
    \end{itemize}}
\end{document}
