\documentclass[openany,letterpaper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}
\usepackage{tgadventor}
\usepackage{XCharter}
\usepackage{titling}
\usepackage{FiraMono}
\usepackage{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage[margin=.8in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{mdframed}
\usepackage{lastpage}
\pagenumbering{gobble}
% FOOTER

\setlist{nosep}
\author{Connor Duncan}
\title{Resume}
\date{\today}
\setlength\parindent{0pt}

% new title command
\renewcommand{\maketitle}{\begin{minipage}[t]{\textwidth}
    \par{\centering{\LARGE{\theauthor}}\par}
    \centering\small\color{gray}

      \underline{\href{https://www.connorduncan.xyz/}{website}} | \underline{\href{https://github.com/ctdunc}{github}} | \href{mailto:connor@connorduncan.xyz}{connor@connorduncan.xyz}
    \end{minipage}
}
% changing title formatting
\titleformat{\section}{\sffamily\scshape\raggedright}{}{0em}{}[{\titlerule[1pt]}]
\titlespacing{\section}{0pt}{3pt}{7pt}
\titleformat{\subsection}{\bfseries}{}{2em}{}
\titlespacing{\subsection}{0pt}{7pt}{7pt}

% itemize
\renewcommand\leftmargini{1.5em}
\renewcommand\labelitemi{$\diamond$}
\newcommand{\Cal}{
    University of California, Berkeley}

\newcommand{\entry}[4]{ % DATE, POSITION, BUSINESS, DUTIES.
    \begin{minipage}[t]{0.15\textwidth}
        \scshape#1
    \end{minipage}
    \vline\hspace{1em}
    \begin{minipage}[t]{0.83\textwidth}
        \raggedright
        {\large\scshape#3}\\
        \vspace{0.3em}
        \vbox{\normalsize\textit{#2}#4}
    \end{minipage}
}

\newcommand{\skill}[2]{ % TYPE, LIST
    \begin{minipage}[t]{0.15\textwidth}
        \scshape#1
    \end{minipage}
    \vline\hspace{1em}
    \begin{minipage}[t]{0.83\textwidth}
        \raggedright
        #2
    \end{minipage}
}
\begin{document}
\maketitle
\vspace{1em}

{\centering Motivated, eager and curious Quantitative Analyst specializing in Fixed-Income ETF portfolio
management across Treasuries, IG/HY/EM Credit, and Municipal Bonds. Math, Linux and Classical/Jazz Piano enjoyer.

}

\section{Education}
\entry{2022--2023}{M.Sc. in Applied Mathematics}{University of Illinois at Chicago}{ || 3.86/4.0 GPA}

\entry{2017--2021}{B.A. in Physics; B.A. in Applied Mathematics}{\Cal}{ || 3.366/4.0 GPA}

\section{Experience}
\entry{2023 --}{Quantitative Analyst (Fixed Income ETF Portfolio Management)}{Invesco}{
    \begin{itemize}
        \item Developed an interactive, introspectable In-Kind Basket Generation/Portfolio Construction tool, decreasing
            decreasing basket negotiation times to as short as 5 minutes.
        \item Implemented a position reconciliation tool, notably improving error detection rates and resolution times with data management.
        \item Developed editor plugins for neovim and VS Code adding zero-overhead syntax highlighting and LSP support for
            Python files containing other languages as string literals (\underline{\href{https://www.connorduncan.xyz/blog/dash-clientside-lsp.html}{blog post}}),
            enabling significant performance and correctness gains.
        \item Prototyped in-house optimizer, using Branch \& Bound to minimize continuous and integer objectives
            with linear/quadratic continuous/integer constraints.
        \item Maintained intra-day flow reconciliation tool ensuring synchronization of basket/trade quantities between our 
            custodian, IBOR and APs.
        \item Managed ongoing transition of front-office systems from Aladdin to Charles River,
            requiring complete re-sourcing and validation of all data ($\sim80$ tables).
        \item Migrated DataFrame libraries from Pandas to Polars, enabling anywhere from $2\times$-$100\times$ performance improvements
            across our ETL.
        \item Spearheaded addition of static type analysis and documentation to critical code.
    \end{itemize}
}

\entry{2022 -- 2023}{Teaching Assistant (Algebra \& Business Linear Algebra)}{University of Illinois at Chicago}{
} 

\entry{Summer 2020}{SULI Intern}
{National Renewable Energy Laboratory}
{
    \begin{itemize}
        \item Developed React app visualizing environmental policy changes in the LA area.
    \end{itemize}
}

\entry{2018--2020}{Research Assistant, Pyle Group}{\Cal; SLAC}{
    \begin{itemize}
        \item Administered CentOS Linux Server, developed readout infrastructure for Dark Matter Search detector R\&D.
    \end{itemize}
}
\section{Technical Skills}
\skill{Python}{
    \begin{itemize}
        \item Polars, SciPy, Flask, Dash, Jinja, threading/multiprocessing, 
            Pydantic, cython.
    \end{itemize}
}
\skill{Linux}{
    \begin{itemize}
        \item Self-hosted websites using both apache and nginx, as well as a mail server.
        \item Daily Arch Linux user since 2018, more comfortable in tmux than Windows.
    \end{itemize}
}

\skill{Other}{
    \begin{itemize}
        \item  By expertise ($\downarrow$): (n)vim, SQL, git, \LaTeX, JavaScript, Bloomberg, Atlassian, GitHub, 
            Rust, Aladdin, MATLAB, lua, Axioma, C/C++, Redis, R, Excel.
    \end{itemize}
}

\section{/etc}
\skill{Interests}{
    \begin{itemize}
        \item {\scshape At Work}: 
            Mixed Integer Programming, Linux, Data Science, Dev Tools, Numerical Analysis of PDEs.
        \item {\scshape After Work}: Piano, Travel, Science Fiction, \underline{
                \href{https://connorduncan.xyz/blog}{blogging}}.\footnote{\underline{\href{https://github.com/ctdunc/ctdunc.github.io/blob/master/site/blog/index.md}{This link}} will work if my site was blocked by your firewall. The .com domain is taken!}
    \end{itemize}}
\end{document}
